Leadership across cultures and contexts: an exploratory analysis

### 2020-04-15 leadership_across_cultures_contexts.Rmd
Lots of edits, esp to text analysis

We then used several methods to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, non-negative matrix factorization (NMF), and hierarchical cluster analysis. Logistic PCA generalizes Pearson's original formulation of PCA to find a projection of the natural parameters of the saturated model that minimizes the Bernoulli deviance [@logisticPCA]. NMF [@lee1999] finds a low rank approximation of the original matrix, $X \approx WH$, where $X$ is an $n \times p$ matrix, and $W$ and $H$ are $n \times k$ and $k \times p$ matrices, respectively; all elements of $X$, $W$ and $H$ are $\geq 0$; and, typically, $k << min(n,p)$. NMF estimates $W$ and $H$ by minimizing a measure of distance (typically the Frobenius norm or the Kullback-Leibler divergence) between $X$ and $WH$ [@gaujoux2010]. Agglomerative hierarchical cluster starts with all elements in their own cluster, and then uses a distance metric to combine the two closest clusters into a single cluster; the next two closest clusters are then combined, and so forth, until all elements belong to a single cluster. We computed distance between our binary vectors using $1-cor(x, y)$, and agglomerated using the Ward algorithm. We assessed the robustness of the clusters using the *pvclust* package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples. We used 10,000 bootstrapped samples.

@@

Our qualities and functions data comprised sparse $`r nrow(leader_text)` \times `r length(quality_vars)`$ and $`r nrow(leader_text)` \times `r length(function_vars)`$ matrices of binary data, respectively.

@@

@@

We used lasso regression as a variable selection technique, which is especially useful when the number of predictors is much greater than the number of cases (i.e., $p > > n$). Lasso regression has a penalty or shrinkage term, $\lambda$: larger values of $\lambda$ retain more variables in the model, at the risk of over-fitting, whereas smaller values of $\lambda$ decrease the number of predictors retained in the model, at the risk of under-fitting. The optimal value of $\lambda$ can be estimated with cross-validation [@tibshirani_regression_1996; @zou_regularization_2005]. We used the glmnet package [@friedman_regularization_2010] to fit a lasso regression model to determine which words best predicted evidence for high status as a leader quality. We used tenfold cross-validation to choose the $\lambda$ parameter that minimized deviance. We then plotted all the non-zero coefficients (Figure \@ref(fig:highstatusplot).

@@

```{r features, fig.width=10, fig.cap="Marginal effects of Poission mixed effects regression models of Feature variables by mode of subsistence. Only models whose subsistence coefficient had a corrected p-value < 0.05. Base level: agriculturalists. See text for details."}

x <- pmap(sub_models_sig, ~ plot(..6, main = ..1))

>>>>

We then used several methods to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, non-negative matrix factorization (NMF), and hierarchical cluster analysis. Logistic PCA generalizes Pearson's original formulation of PCA to find a projection of the natural parameters of the saturated model that minimizes the Bernoulli deviance [@logisticPCA].

NMF [@lee1999] finds a low rank approximation of the original matrix, $X \approx WH$, where $X$ is an $n \times p$ matrix, and $W$ and $H$ are $n \times k$ and $k \times p$ matrices, respectively; all elements of $X$, $W$ and $H$ are $\geq 0$; and, typically, $k << min(n,p)$. NMF estimates $W$ and $H$ by minimizing a measure of distance between $X$ and $WH$ [we used the default "brunet" algorithm, which uses the Kullback-Leibler divergence; @gaujoux2010].

Agglomerative hierarchical cluster analysis starts with all elements in their own cluster, and then uses a distance metric to combine the two closest clusters into a single cluster; the next two closest clusters are then combined, and so forth, until all elements belong to a single cluster. We computed distance between our binary vectors using $1-cor(x, y)$, and agglomerated using the Ward algorithm. We assessed the robustness of the clusters using the *pvclust* package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples. We used 10,000 bootstrapped samples.

@@

```{r}
num_removed <- sum(rowSums(leader_text2[all_study_vars]) == 0)
```

To investigate features in our data, we first removed `r num_removed` text records with no evidence for any variable.

Our qualities and functions data comprised sparse $`r nrow(leader_text2)` \times `r length(quality_vars)`$ and $`r nrow(leader_text2)` \times `r length(function_vars)`$ matrices of binary data, respectively.

@@

## Ethnographer and Culture random effects

```{r FunctionRE, fig.cap="Ethnographer and Culture random effects for function variables."}
plot_fun_auth_cult_re
```

```{r QualityRE, fig.height=8, fig.cap="Ethnographer and Culture random effects for quality variables."}
plot_qual_auth_cult_re
```

@@

We used lasso regression as a variable selection technique, which is especially useful when the number of predictors is much greater than the number of cases (i.e., $p > > n$). Lasso regression has a penalty or shrinkage term, $\lambda$: larger values of $\lambda$ retain more variables in the model, at the risk of over-fitting, whereas smaller values of $\lambda$ decrease the number of predictors retained in the model, at the risk of under-fitting. The optimal value of $\lambda$ can be estimated with cross-validation [@tibshirani_regression_1996; @zou_regularization_2005]. We used the glmnet package [@friedman_regularization_2010] to fit a lasso regression model to determine which words best predicted evidence for high status as a leader quality. We used tenfold cross-validation to choose the $\lambda$ parameter that minimized deviance. We then plotted all the non-zero coefficients (Figure \@ref(fig:highstatusplot)).

@@

```{r features, fig.width=8, fig.cap="Esimtated marginal means [@lenth2020] of Poission mixed effects regression models of Feature variables by mode of subsistence. Only models whose subsistence coefficient had a corrected p-value < 0.05. Plotted on the response scale. See text for details."}

# x <- pmap(sub_models_sig, ~ plot(..6, main = ..1))

plot_feature_models
